Score-based tests of measurement invariance: Use in practice

### 2013-09-17 frontiers/frontiers.Rnw
Work on frontiers paper: o Small revisions to paper o Small revisions/additions to references o Added LMuo to sctest.lavaan
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@261 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\abstract{
We study a family of 
recently-proposed measurement invariance tests that are derived from
stochastic processes.  As applied to factor analysis and other
psychometric models, the tests are advantageous in that (i) they do
not require advance definition of subgroups of cases that violate
measurement invariance; (ii) they can potentially identify specific
model parameters that violate measurement invariance; and (iii) they
do not require a model of additional complexity to be estimated.  
Because the tests have been applied to only a small number of
psychometric examples, we conduct a series of simulation studies that provide
a detailed examination of test properties.  These studies allow us
to study (details here).  The studies indicate that.}

\SweaveOpts{engine = R, eps = FALSE, echo = FALSE, pdf=TRUE}

\acknowledgements{This work was supported by National Science
  Foundation grant SES-1061334. 
  Correspondence to. 
  Email: }
\shorttitle{Measurement invariance test properties}
\rightheader{Measurement invariance test properties}

\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
%% for internal use
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}
\newcommand{\seeme}[1]{\emph{\marginpar{SEEME} (#1)}}

@@

Parameter invariance is a fundamental assumption for measurement model in 
developing psychometric scales. Violation of it could happen in any estimated parameter.  
In order to fix the psychometric scale, we need to specify the violating parameter.  
Moreover, in practice model misspecification is an inevitable issue to a certain extent 
and may interfere with parameter invariance specificity. A family of tests derived from 
stochastic process has been recently proposed to identify specific parameters impacted 
by measurement invariance violation. However, these tests have only been used to a limited 
type of parameter violation. It is this paper's intent to provide a detailed examination 
of the tests' performance under practical research scenarios.

@@

Following model estimation, the tests primarily work on the {\em
  scores} of the fitted model; these are defined as
%% TODO indicate that this is evaluated at theta hat
\begin{equation}
\label{eq:score}
s(\hat{\theta};x_{i}) = \left ( \frac{\partial
      \ell(\theta;x_{i})}{\partial(\theta_{1})},...,\frac{\partial
      \ell(\theta;x_{i})}{\partial(\theta_{k})} \right )^{T},
\end{equation} 
where $\ell(\theta; x_i)$ is the likelihood associated with individual
$i$, $\theta$ is a $k$-dimensional parameter vector, and
$\hat{\theta}$ is the maximum likelihood estimate.  The cross-product
of these scores forms the ``meat'' for the calculation of robust
(Huber-White) standard errors \cite{freed06}. \seeme{not sure about this citation}

@@

For an ordinal auxiliary variable with $m$ levels, these modifications are 
based on $t_{\ell}$ $(\ell=1,\ldots,m-1)$, which are the empirical, cumulative 
proportions of individuals observed at the first $m-1$ levels.
The modified statistics are then given by
\begin{eqnarray}
    \label{eq:wdm}
      \mathit{WDM}_o & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~ \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1/2}      
                             \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} |,\\
    \label{eq:maxlmo}
      \max \mathit{LM}_o & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~
                             \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
                             \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2,
\end{eqnarray}
where $i_{\ell}=\lfloor n \cdot t_{\ell} \rfloor$ $(\ell=1,\ldots,m-1)$.
Critical values associated with the $\mathit{WDM}_o$ statistic can be obtained 
directly from a multivariate normal distribution \cite<see>{hotzei08}, while 
critical values associated with $\max \mathit{LM}_o$ can be obtained via simulation.  
This simulation is somewhat computationally intensive and, in practice, takes about 
10 minutes on the authors' computers.  However, the wait appears to be worth it, 
as \citeA{MerFanZei} found the performance of the $\max \mathit{LM}_o$ statistic 
to be considerably better than the $\mathit{WDM}_o$ statistic.

@@

If $V$ is only nominal/categorical,
there is not even a partial ordering, i.e., measurement invariance tests should neither
exploit the ordering of $V$'s levels nor of the observations within the level.
In this situation, unordered $\mathit{LM}$ is proposed.  This test statistic can be formally written as
\begin{equation}
    \label{eq:lmuo}
     \mathit{LM}_\mathit{uo} = \sum_{\ell = 1, \dots, m} \sum_{j = 1, \dots, k}
       \left( {\bm B}(\hat {\bm \theta})_{i_\ell j} - {\bm B}(\hat {\bm \theta})_{i_{\ell - 1}j} \right)^2,
\end{equation}
This test statistic discards the ordinal nature of the auxiliary
variable, essentially similar to the nature of likelihood ratio tests (or,
equivalently, via Wald tests or Lagrange multiplier tests): the ordinality of the auxiliary variable is
ignored. By contrast, this statistics is easier to get compared to likelihood ratio test statistics since it doesn't involve specifying different sets of competitive models.

\SweaveOpts{engine = R, eps = FALSE, echo = TRUE, pdf=TRUE}
\section{Tutorial}

@@

Another similar statistics is $\mathit{LM}_\mathit{uo}$ as shown in Equation~ \eqref{eq:lmuo}. We can obtain this statistic by using the following function. 
<<LM_UO>>=
## unordered LM
lm_u0 <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="range")
round(c(lm_u0$p.value), 2)
@
we wish to test model parameter numbers 1 through 4 (which are the factor loadings), 
which are supplied by lavaan.  The ordering of parameters can be obtained via the 
command \verb+coef(reduced)+.

Because our sample size is large, this likelihood ratio test is sensitive to 
small measurement invariance violations \cite{Ben80}.  Thus, the above test may 
be significant solely due to the large sample size.  Additionally, the above test 
is sensitive to small measurement invariance violations that are not likely to be 
of interest to researchers.  For example, imagine that the 15-year-olds' 
parameters are slightly different than the other age groups.  
The 15-year-olds are in the middle of the age groups, and there 
is not likely to be any theoretical justification for 15-year-olds 
differing from every other age group.  We need test statistics that 
are sensitive to measurement invariance violations that are monotonic 
with age.  The score-based statistics described previously possess this property.

\subsection{Continuous Treatment}
If we are interested in measurement invariance violations that are 
monotonic with the age groups, it is perhaps simplest to treat the 
age groups as continuous.  In doing so, we can use the statistics 
from Equations~\eqref{eq:dmax},~\eqref{eq:cvm}, and~\eqref{eq:maxlm}.  
That is, we can fit a model whose parameters 
are restricted to be equal across all individuals and then examine 
how individuals' scores $s(\hat{theta};x_{i})$ fluctuate with their
age (where age ties are broken arbitrarily).
This is demonstrated below, with similar code being useful when one is 
testing measurement invariance w.r.t.\ other continuous variables.

@@

To proceed, we focus only on the restricted model that was fit in 
the previous section (the object \verb+reduced+).  We obtain scores 
associated with the factor loadings from this model, then test the 
fluctuation of these scores across \verb+agegroup+.  
As demonstrated below, the score computation and testing can be efficiently carried out using \verb+sctest.lavaan()+.
%% Ting: I removed this because the model was already fit in the
%%       previous section
%%<<theta estimate>>=
%%theta<-cfa(
%%  'f1=~gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5',
%%  data=yg,group="agegroup",meanstructure=TRUE,
%%  group.equal="loadings")
%%@

<<Get Estimate>>=
dm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="DM")
cvm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="CvM")
maxlm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLM")
round(c(dm$p.value, cvm$p.value, maxlm$p.value), 2)
@
%% move this code description to the cat2BB part. 
%In the code above, the \verb+parm=1:4+ argument means that 
%we wish to test model parameter numbers 1 through 4 (which are the factor loadings), 
%which are supplied by lavaan.  The ordering of parameters can be obtained via the 
%command \verb+coef(reduced)+.

We see that two of the three $p$-values output at 
the end of the code are larger than that associated with the LRT (with the CvM statistic being non-significant).

@@

%% <<conti3resmaxLM>>=
%% maxLM<-sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLM", fplot=FALSE) 
%% maxLM
%% @
The tests carried out here assume a unique ordering of individuals by age, but 
this is obviously not the case.  To compute the statistics and $p$-values, 
the strucchange package implicitly assumed an arbitrary ordering of 
individuals who are tied on age.  If we were to change this ordering, 
the resulting statistics and $p$-values would also change, which is obviously problematic.  
To accurately account for the multiple observations  
at the same age level, we must used the ordinal tests from 
Equations~\eqref{eq:wdm} and~\eqref{eq:maxlmo}.  These are described next.

@@

%% As a matter of fact, there is only a partial ordering of individuals 
%% with respect to $V$,i.e. observations with the same level 
%% of $V$ have no unique ordering. The ordinal statistics 
%% proposed here are similar to those described in Equation 
%% and above, except that we focus on "bins" of individuals 
%% at each level of the ordinal variable.

The difference between the ordinal test statistics and their
continuous counterparts is that the ordinal statistics do not require
an ordering of individuals  
within the same age group.  To compute the test statistics, we allow
the scores of all tied individuals to enter the cumulative sum
(Equation~\eqref{eq:cumscore}) simultaneously.  This results in
modified critical values and test statistics that are sensitive to
measurement invariance violations that are monotonic with age group.
%% Therefore, this ordinal test should 
%% reflect more real status of the data. In terms of code, we still need the gefp 
%% function to get the B estimate. After that, we can use different 
%% functions to aggregate and scale B estimation. 
%% The function are called ordwmax and ol2bb1.fun, respectively.

To carry out the tests, we can rely on the same function that we used
for the continuous test statistics.  As mentioned previously,
calculation of the ``maxLMo'' statistic (Equation~\eqref{eq:maxlmo})
can be lengthy from the need to simulate critical values.

<<critvals,echo=FALSE>>=
if(file.exists("opval1.rda")){
  load("opval1.rda")
} else {
  opval1 <- ordL2BB(gefp1)
  save(opval1, file="opval1.rda")
}
## Get test statistics using fnl=opval1 without showing it in the paper
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo",
                 fnl=opval1)
@

<<ordtest,eval=FALSE>>=
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo")
@ 
<<pvals>>=
round(c(wdmo$p.value, maxlmo$p.value), 2)
@

\seeme{I don't know how to adjust the line interval space for the code part.}

In computing the ordinal test statistics, we
obtain~$p=\Sexpr{round(wdmo$p.value,2)}$
and~$p=\Sexpr{round(maxlmo$p.value,2)}$, respectively.
Both $p-$values are clearly larger than that of the 
likelihood ratio test and neither is significant at $\alpha=0.05$. 
This provides evidence that there is no measurement
invariance violation that is monotonic with age group.  Instead, given
the large sample size, the likelihood ratio test may be overly
sensitive to anomalous, non-monotonic violations at one (or a few) age groups.

Besides, plots can be generated by using \verb+fplot=TRUE+ argument inside of \verb+sctest.lavaan()+ function. Plots representing the statistics' fluctuations across levels of age group are displayed in Figure~\ref{fig:ordres}. The left panel displays the process associated with $\mathit{WDM}$. The right panel displays the process associated with $\mathit{maxLM}$. In both cases, the test statistics in the sequence assess a split of the observations up to age group $i$ vs. greater than $i$, and the null hypothesis is rejected if the maximum of the statistics is larger than its 5\% critical value(visualized by the horizontal red line). The final group is not displayed because it encompass all observations and always equals to zero. The age group 1:5 in the x-axis corresponds to 10-11, 12-13, 14,15,16 respectively. It is observed that both statistics generally increase with age.

@@

%Or alternatively, we could use the sctest.lavaan function, but assign ordinal statistics.
\begin{figure}[H]
\caption{Fluctuation processes for the WDM statistic(left panel) and the maxLM statistic(right panel)}
\label{fig:ordres}
\setkeys{Gin}{width=\textwidth}
<<ordres, fig=TRUE, height=3, width=6, echo=FALSE>>=
par(mfrow = c(1, 2), cex = 0.65)
WDMo<-sctest(reduced, as.numeric(yg$agegroup),parm=1:4, stat="WDMo", fplot=TRUE,ylim=c(0,3.5),xlab="Age group",ylab = "Weighted max statistics")

#WDMo 
maxLMo<-sctest(reduced,as.numeric(yg$agegroup),parm=1:4, stat="maxLMo", fnl=opval1, fplot=TRUE,ylim=c(0,14),xlab="Age group",ylab = "LM statistics")
#maxLMo 
@
\end{figure}

\seeme{can't change age group axis, since order.by argument needs agegroup as numeric}

@@

\SweaveOpts{engine = R, eps = FALSE, echo = FALSE, pdf=TRUE}
\section{Simulation}
\section{Simulation 1}
In Simulation 1, we examined the extent to which the proposed tests
are sensitive to the specific model parameter that violates measurement
invariance.  If, say, a factor loading violates measurement
invariance, it is plausible that this violation impacts other
parameters such as factor covariances or other loadings 
associated with the factor in question.  Thus, the goal of the
simulation was to examine the extent to which the proposed tests
attribute the measurement invariance violation to the parameters that
are truly in violation.

@@

<<sim1>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

@@

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("maxLM_O","WDM_O","LM_uO")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("maxLM_o","WDM_o","LM_uo"))
  levels(sim1$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")

>>>>

\abstract{
In this paper, we consider a family of 
recently-proposed measurement invariance tests that are based on the {\em scores} of a fitted model.  This family can be used to test for measurement invariance
w.r.t.\ a continuous auxiliary variable, without pre-specification of subgroups.
Moreover, the family can be used when one wishes to test for 
measurement invariance w.r.t.\ an ordinal auxiliary variable, yielding test
statistics that are sensitive to violations that are monotonically related to
the ordinal variable (and less sensitive to non-monotonic violations).  The paper is specifically aimed at potential users of the tests who may wish to know (i) how the tests can be employed for their data, and (ii) whether the tests can accurately identify measurement invariance violations in the presence of model misspecification.  After providing an overview of the tests, we illustrate their general use via the R packages lavaan and strucchange.  We then describe two novel simulations that provide evidence of the tests' practical abilities.  As a whole, the paper provides researchers with the tools and knowledge needed to apply these tests to general measurement invariance scenarios.}

\acknowledgements{This work was supported by National Science
  Foundation grant SES-1061334. 
  Correspondence to Edgar Merkle. 
  Email: merklee@missouri.edu.}
\shorttitle{Measurement invariance test properties}
\rightheader{Measurement invariance test properties}

\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
%% for internal use
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\readme}[1]{\emph{\marginpar{Ed} (#1)}}
\newcommand{\seeme}[1]{\emph{\marginpar{Ting} (#1)}}

@@

\readme{We need a better intro here.  The special topic is about measurement invariance, so most of this paragraph is obvious to everyone.  Maybe start with how so many other papers are about BSEM.}
Parameter invariance is a fundamental assumption for measurement model in 
developing psychometric scales. Violation of it could happen in any estimated parameter.  
In order to fix the psychometric scale, we need to specify the violating parameter.  
Moreover, in practice model misspecification is an inevitable issue to a certain extent 
and may interfere with parameter invariance specificity. A family of tests derived from 
stochastic process has been recently proposed to identify specific parameters impacted 
by measurement invariance violation. However, these tests have only been used to a limited 
type of parameter violation. It is this paper's intent to provide a detailed examination 
of the tests' performance under practical research scenarios.

@@

Following model estimation, the tests primarily work on the {\em
  scores} of the fitted model; these are defined as
%% TODO indicate that this is evaluated at theta hat
\begin{equation}
\label{eq:score}
s(\hat{\theta};x_{i}) = \left ( \frac{\partial
      \ell(\theta;x_{i})}{\partial(\theta_{1})},...,\frac{\partial
      \ell(\theta;x_{i})}{\partial(\theta_{k})} \right )^{T},
\end{equation} 
where $\ell(\theta; x_i)$ is the likelihood associated with individual
$i$, $\theta$ is a $k$-dimensional parameter vector, and
$\hat{\theta}$ is the maximum likelihood estimate.  The cross-product
of these scores forms the ``meat'' for the calculation of robust
(Huber-White) standard errors \cite<e.g.,>{zei06b}.

@@

For an ordinal auxiliary variable with $m$ levels, these modifications are 
based on $t_{\ell}$ $(\ell=1,\ldots,m-1)$, which are the empirical, cumulative 
proportions of individuals observed at the first $m-1$ levels.
The modified statistics are then given by
\begin{eqnarray}
    \label{eq:wdm}
      \mathit{WDM}_o & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~ \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1/2}      
                             \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} |,\\
    \label{eq:maxlmo}
      \max \mathit{LM}_o & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~
                             \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
                             \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2,
\end{eqnarray}
where $i_{\ell}=\lfloor n \cdot t_{\ell} \rfloor$ $(\ell=1,\ldots,m-1)$.
Critical values associated with the $\mathit{WDM}_o$ statistic can be obtained 
directly from a multivariate normal distribution \cite<see>{hotzei08}, while 
critical values associated with $\max \mathit{LM}_o$ can be obtained via simulation.  
This simulation is somewhat computationally intensive and, in practice, takes about 
10 minutes on the authors' computers.  However, the wait is often worth it, 
as \citeA{MerFanZei} found the performance of the $\max \mathit{LM}_o$ statistic 
to be considerably better than the $\mathit{WDM}_o$ statistic.

@@

Finally, if $V$ is only nominal/categorical, the cumulative sums of scores can be used to obtain the a Lagrange multiplier statistic.
This test statistic can be formally written as
\begin{equation}
    \label{eq:lmuo}
     \mathit{LM}_\mathit{uo} = \sum_{\ell = 1, \dots, m} \sum_{j = 1, \dots, k}
       \left( {\bm B}(\hat {\bm \theta})_{i_\ell j} - {\bm B}(\hat {\bm \theta})_{i_{\ell - 1}j} \right)^2,
\end{equation}
and it is asymptotically equivalent to the usual, likelihood ratio test statistic.  It is advantageous over the likelihood ratio test because it requires estimation of only one model (the restricted model).  We make use of this advantage in the simulations, described later.

\section{Tutorial}

@@

\readme{I just added LMuo to the sctest command; the range statistic is different.}
Another similar statistics is $\mathit{LM}_\mathit{uo}$ as shown in Equation~ \eqref{eq:lmuo}. We can obtain this statistic by using the following function. 
<<LM_UO>>=
## unordered LM
lm_uo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="LMuo")
round(lm_uo$p.value, 4)
@
The above command specifies that we wish to test model parameter numbers 1 through 4, which are the factor loadings supplied by lavaan.  The ordering of parameters can be obtained via the command \verb+coef(reduced)+.

Because our sample size is large, this likelihood ratio test is known to be sensitive to 
small measurement invariance violations \cite{Ben80}.  That is, the above tests 
are sensitive to small measurement invariance violations that are not likely to be 
of interest to researchers.  For example, imagine that the 15-year-olds' 
parameters are slightly different than the other age groups.  
The 15-year-olds are in the middle of the age groups, and there 
is not likely to be any theoretical justification for 15-year-olds 
differing from every other age group.  One solution to this problem would be the {\em approximate invariance} methods proposed by Muth\'{e}n and others.\readme{Ting: Can you get some approximate invariance references and put them here?}
Alternatively, we can use the proposed family of score-based statistics to obtain tests that are sensitive to the ordering of age.

\subsection{Continuous Treatment}
If we are interested in measurement invariance violations that are 
monotonic with the age groups, it is perhaps simplest to treat the 
age groups as continuous.  In doing so, we can use the statistics 
from Equations~\eqref{eq:dmax},~\eqref{eq:cvm}, and~\eqref{eq:maxlm}.  
That is, we can fit a model whose parameters 
are restricted to be equal across all individuals and then examine 
how individuals' scores $s(\hat{\theta};x_{i})$ fluctuate with their
age (where age ties are broken arbitrarily).
This is demonstrated below, with similar code being useful when one is 
testing measurement invariance w.r.t.\ truly continuous variables.

@@

To proceed, we focus only on the restricted model that was fit in 
the previous section (stored in the object \verb+reduced+).  We obtain scores 
associated with the factor loadings from this model, then test the 
fluctuation of these scores across \verb+agegroup+.  
As demonstrated below, the score computation and testing can be efficiently carried out using \verb+sctest.lavaan()+.
%% Ting: I removed this because the model was already fit in the
%%       previous section
%%<<theta estimate>>=
%%theta<-cfa(
%%  'f1=~gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5',
%%  data=yg,group="agegroup",meanstructure=TRUE,
%%  group.equal="loadings")
%%@

<<Get Estimate>>=
dm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="DM")
cvm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="CvM")
maxlm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLM")
round(c(dm$p.value, cvm$p.value, maxlm$p.value), 2)
@
%% move this code description to the cat2BB part. 
%In the code above, the \verb+parm=1:4+ argument means that 
%we wish to test model parameter numbers 1 through 4 (which are the factor loadings), 
%which are supplied by lavaan.  The ordering of parameters can be obtained via the 
%command \verb+coef(reduced)+.  
We see that two of the three $p$-values output at 
the end of the code are larger than that associated with the LRT (with the CvM statistic being non-significant).

@@

%% <<conti3resmaxLM>>=
%% maxLM<-sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLM", fplot=FALSE) 
%% maxLM
%% @
The tests carried out here assume a unique ordering of individuals by age, but 
this is obviously not the case.  To compute the statistics and $p$-values, 
the strucchange package implicitly assumed an arbitrary ordering of 
individuals who are tied on age.  If we were to change this ordering, 
the resulting statistics and $p$-values would also change, which is obviously problematic.  
To accurately account for the multiple observations  
at the same age level, we must use the ordinal tests from 
Equations~\eqref{eq:wdm} and~\eqref{eq:maxlmo}.  These are described next.

@@

%% As a matter of fact, there is only a partial ordering of individuals 
%% with respect to $V$,i.e. observations with the same level 
%% of $V$ have no unique ordering. The ordinal statistics 
%% proposed here are similar to those described in Equation 
%% and above, except that we focus on "bins" of individuals 
%% at each level of the ordinal variable.  
The main difference between the ordinal test statistics and their
continuous counterparts is that the ordinal statistics do not require
an ordering of individuals  
within the same age group.  To compute the test statistics, we allow
the scores of all tied individuals to enter the cumulative sum
(Equation~\eqref{eq:cumscore}) simultaneously.  This results in
modified critical values and test statistics that are sensitive to
measurement invariance violations that are monotonic with age group.
%% Therefore, this ordinal test should 
%% reflect more real status of the data. In terms of code, we still need the gefp 
%% function to get the B estimate. After that, we can use different 
%% functions to aggregate and scale B estimation. 
%% The function are called ordwmax and ol2bb1.fun, respectively.

To carry out the tests, we can rely on the same function that we used
for the continuous test statistics.  As mentioned previously,
calculation of the ``maxLMo'' statistic (Equation~\eqref{eq:maxlmo})
can be lengthy from the need to simulate critical values.
%% This is more complicated than necessary, but it is written like
%% this so that we can load the ordinal critical values behind the
%% scenes without confusing the reader.
<<critvals,echo=FALSE,strip.white=all>>=
if(file.exists("opval1.rda")){
  load("opval1.rda")
} else {
  opval1 <- ordL2BB(gefp1)
  save(opval1, file="opval1.rda")
}
## Get test statistics using fnl=opval1 without showing it in the paper
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo",
                 fnl=opval1)
@ 
<<ordtest,eval=FALSE,strip.white=all>>=
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo")
@ 
<<pval,strip.white=all>>=
round(c(wdmo$p.value, maxlmo$p.value), 2)
@ 
In computing the ordinal test statistics, we
obtain~$p=\Sexpr{round(wdmo$p.value,2)}$
and~$p=\Sexpr{round(maxlmo$p.value,2)}$, respectively.
Both $p-$values are clearly larger than that of the 
likelihood ratio test and neither is significant at $\alpha=0.05$. 
This provides evidence that there is no measurement
invariance violation that is monotonic with age group.  Instead, given
the large sample size, the likelihood ratio test may be overly
sensitive to anomalous, non-monotonic violations at one (or a few) age groups.

In addition to test statistics, ``instability plots'' can be generated by using the \verb+fplot=TRUE+ argument within \verb+sctest.lavaan()+. Plots representing the statistics' fluctuations across levels of age group are displayed in Figure~\ref{fig:ordres}. The left panel displays the process associated with $\mathit{WDM}_o$, and the right panel displays the process associated with $\max\mathit{LM}_o$. 
%The following sentence is not correct for all statistics.  
%In both cases, the test statistics in the sequence assess a split of the observations up to age group $i$ vs. greater than $i$, and the null hypothesis is rejected if the maximum of the statistics is larger than its 5\% critical value(visualized by the horizontal red line). 
These plots display the loadings' fluctuation across age groups, with the x-axis reflecting age group and the y-axis reflecting test statistic values (larger values reflect more instability).  The red lines reflect critical values, so that the hypothesis of measurement invariance is rejected if a black line crosses the red line.  While the measurement invariance tests are non-significant, the plots imply some instability in the older age groups (15 and 16).

@@

%Or alternatively, we could use the sctest.lavaan function, but assign ordinal statistics.
\begin{figure}[H]
\caption{Fluctuation processes for the $\mathit{WDM}_o$ statistic (left panel) and the $\max\mathit{LM}_o$ statistic (right panel).}
\label{fig:ordres}
\setkeys{Gin}{width=\textwidth}
<<ordres, fig=TRUE, height=3, width=6, echo=FALSE, results=hide>>=
par(mfrow = c(1, 2), cex = 0.65)
sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo", 
       fplot=TRUE, ylim=c(0,3.5), xlab="Age group",
       ylab = "Weighted max statistics", axes=FALSE)
axis(1, at = 1:5, labels = c("10-11", "12-13", "14", "15", "16"))
box()

sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo", 
       fnl=opval1, fplot=TRUE, ylim=c(0,14), xlab="Age group",
       ylab = "LM statistics", axes = FALSE)
axis(1, at = 1:5, labels = c("10-11", "12-13", "14", "15", "16"))
box()
@
\end{figure}

@@

\section{Simulation 1}
In Simulation 1, we examined the extent to which the proposed tests
are sensitive to the specific model parameter that violates measurement
invariance.  If, say, a factor loading violates measurement
invariance, it is plausible that this violation impacts other
parameters such as factor covariances or other loadings 
associated with the factor in question.  Thus, the goal of the
simulation was to examine the extent to which the proposed tests
attribute the measurement invariance violation to the parameters that
are truly in violation.

@@

<<sim1,echo=FALSE>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

@@

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("maxLM_O","WDM_O","LM_uO")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("maxLM_o","WDM_o","LM_uo"))
  levels(sim1$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")  
@
